Cartogram Mapping and Application to Cancer Data.

### 2019-06-13 review_article.Rmd
draft of choropleth development and use for cancer mapping

Cartographer lit, cancer atlas lit

@MACM
geographically mapping cancer mortality uses maps of europe:
Static choropleth maps, coloured from green to red, supplemented by histograms.
Introduces Moran scatterplot "In the Moran’s scatterplot, each geographic unit is represented by a point, whose X coordinate is the value of the mortality in that area and the Y coordinate is equal to the average value of the theme in the bordering areas'
And Moran's index: spatial correlation with neighbours

'visual intuition coming from the inspection of maps must be complemented and validated by rigorous statistical analyses.'
Does not acknowledge the map as a visual statistic, but a tool.

>>>>

Choropleth mapping methods explore a relationship across a geographical space, utilising administrative or other boundaries. In thematic maps colours or patterns are used represent levels, or rates, of the information communicated. @CIBMUK believes these 'representations are highly intuitive' as the familiar shapes help those looking at the maps to infer spatial relationships.  
They are true maps, used for visual inspection and can help to direct hypotheses about the statial patterns across the physical space. 
These true maps rely on visual perception, a fallible metric as it can 'weight the shades from the largest areas more heavily.' The inferences derived from visual inspection, and the hypotheses developed can suffer from 'bias'. The use of thematic disease mapping was motivated by a 'desire to evaluate geographical patterns' @DMAHP. This approach has been seen as a 'trigger for questions' which drives research of spatial distributions.  @MACM also recognises the intuition that can be derived from choropleths. However, his critique is that they must be 'validated by rigorous statistical analyses.'

As a cancer mapping tool choropleths are often used. Specifically in epidemiology, the study of the distribution of health-related states and events. Almost 100 years of cancer maps have explored the US and the UK with increased effectiveness in presentation of unbiased rates. @DMAHP cites Cruickshank's (1947) discussion of 'formal statistical assessment of the spatial pattern' as another major advancement.

The data collection methods, and administrative control within regions lends itself to the presentation of cancer mortality rates across regions.
